---
title: "Lab 4 Team"
author: "Meredith Tucker, (mltucker2001) and Mike Zaino, (mzaino16)"
date: "3/11/2020"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(readxl)
library(tidyr)
```


```{r}
defense <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
str(defense)
offense <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
str(offense)
biography <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
str(biography)
```

# Part One
1.) The player names and opponent names are factors.

```{r}
defclean <- defense %>% 
  mutate_at(c('Name', 'Opponent_Opponent'), factor)

offclean <- offense %>% 
  mutate_at(c('Name', 'Opponent_Opponent'), factor)

bioclean <- biography %>% 
  mutate_at(c('Name'), factor)
```


2.) All offensive and defensive statistics are numerical. The Weight column in biogrpahy is also numeric. (Hint: look at ?mutate_at if you want to save writing some code).

```{r}
defclean <- defclean %>% 
  mutate_at(c(3,4,5,6,7,8,9,10,11), as.numeric)
offclean <- offclean %>% 
  mutate_at(c(3,4,5,6,7,8,9,10,11, 12), as.numeric)

head(defclean)
head(offclean)

```


3.) Change the Height column in biogrpahy into numeric. (Hint: Use a unit that works well)
```{r}
bioclean <- bioclean %>% 
  separate(Height, c('Feet', 'Inches'), sep='-', convert=TRUE) %>% mutate(cm = (12*Feet + Inches) *2.54)
head(bioclean)
```

# Part 2

For the following questions, work on the cleaned data frames defClean, offClean, and bioClean.

1.) Reformat the defClean data frame into a tidy format using pivot_longer. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.
```{r}

longdefclean <- defclean %>% 
  pivot_longer(c(3:11), names_to='Statistic', values_to='Score')

```


2.) Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?
```{r}
longdefclean %>% 
  ggplot(aes(x=Score, fill=Statistic)) + geom_histogram() + facet_wrap(~Statistic)
```



3.) Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense.

```{r}
longdefclean %>% 
  filter(Opponent_Opponent == 'West Virginia' | Opponent_Opponent =='Kansas State') %>% 
  filter(Statistic == 'Tackles_Solo') %>% 
  group_by(Name) %>% 
  mutate(Count = n()) %>% 
  filter(Count== 2) %>% 
  ungroup() %>% 
  pivot_wider(names_from = Opponent_Opponent, values_from = Score) %>% 
  ggplot(aes(x=`West Virginia`, y=`Kansas State`)) + geom_point() + stat_smooth(method=lm)
```

West Virginia has more Solo Tackles than Kansas State. Iowa State had a better defense against West Virginia than against Kansas State. 

# Now turn our attention to the player biography.

1.) Separate the Hometown column into two columns, namely the city and the state. Print the first few rows of these two columns. (Hint: look at the sep= argument of separate)

```{r}
bioclean <- bioclean %>% 
  separate(Hometown, c('City', 'State'), sep=',', convert=TRUE)
head(bioclean)
```


2.) How many players are there from each state?
```{r}
bioclean %>%
  group_by(State) %>% 
  summarize(n=n())
```




